Determinism, Complexity, and Predictability in Computer Performance

### 2014-04-15 concl.tex
some fill-ins on concl.tex

In this paper, we have argued that redundancy~\cite{crutchfield2003}
is a useful definition of the inherent predictability of a time
series.  To operationalize that definition, we use an approximation of
the Kolmogorov-Sinai entropy~\cite{lind95}, estimated using a weighted
version of the permutation entropy of~\cite{bandt2002per}.  This WPE
technique---an ordinal calculation of forward information transfer in
a time series---is ideal for our purposes because it works with
real-valued data and is known to converge to the true entropy value.
Using a variety of forecast models and more than a hundred time-series
data sets from a computer-performance experiment, we have shown that
prediction accuracy is indeed correlated with weighted permutation
entropy: the higher the WPE, in general, the higher the prediction
error.  There are some exceptions: [[Liz to do: summarize \& discuss
    here.]]

@@

An important practical corollary to this empirical correlation of
predictability and WPE is a practical strategy for assessing
appropriateness of forecast methods.  If the forecast produced by a
particular method is poor but the time series contains a significant
amount of predictive structure, one can reasonably conclude that that
method is inadequate to the task and that one should seek another
method.  The nonlinear LMA method, for instance, performs better in
most cases because it is more general.  The \naive ~method, which
simply predicts the mean, works better in noisy signals because it
effects a filtering operation.  The simple random-walk strategy is
better than LMA, ARIMA, or the \naive ~method on the \gcc signal,
which is extremely complex---i.e., extremely low redundancy.
{\color{red} I suggest we don't mention \svdone here since it's a
  pretty detailed thing and conclusions are for high-level summaries.
  It's a good point, though, and I'll leave it in comments in the
  file.}
%  \naive ~wins on \svdone with the exception to the 5 outliers.
The dashed line in Figures~\ref{fig:wpe_vs_mase_best}
and~\ref{fig:wpe_vs_mase_all} is a useful empirical guideline for
knowing when a model is not well-matched to the task at hand: a $MASE$
score that is more than {\color{red}[[what's the equation for that
      line, Ryan?]]} times the WPE of the corresponding signal is an
indication that that time series has more predictive structure than
that forecast model can capture and exploit.

>>>>

In this paper, we have argued that redundancy~\cite{crutchfield2003}
is a useful definition of the inherent predictability of a time
series.  To operationalize that definition, we use an approximation of
the Kolmogorov-Sinai entropy~\cite{lind95}, estimated using a weighted
version of the permutation entropy of~\cite{bandt2002per}.  This WPE
technique---an ordinal calculation of forward information transfer in
a time series---is ideal for our purposes because it works with
real-valued data and is known to converge to the true entropy value.
Using a variety of forecast models and more than a hundred time-series
data sets from a computer-performance experiment, we have shown that
prediction accuracy is indeed correlated with weighted permutation
entropy: the higher the WPE, in general, the higher the prediction
error.  There are some exceptions: \svdone is easier to predict than
its WPE value suggests, for instance, using any of the three forecast
strategies studied here.  Conversely, forecast errors for the \col and
\svdfive signals are higher than the corresponding WPEs
suggest---except for the nonlinear LMA predictor.

@@

An important practical corollary to this empirical correlation of
predictability and WPE is a practical strategy for assessing
appropriateness of forecast methods.  If the forecast produced by a
particular method is poor but the time series contains a significant
amount of predictive structure, one can reasonably conclude that that
method is inadequate to the task and that one should seek another
method.  The nonlinear LMA method, for instance, performs better in
most cases because it is more general.  This becomes particularly
apparent in the \col and \svdfive examples, where the other two
methods do not perform well.  The \naive ~method, which simply
predicts the mean, works better in noisy signals because it effects a
filtering operation.  The simple random-walk strategy outperforms
nLMA, ARIMA, and the \naive ~method on the \gcc signal, which is
extremely complex---i.e., extremely low redundancy.  {\color{red} I
  suggest we don't mention \svdone here since getting into outliers is
  a pretty detailed thing and conclusions are for high-level
  summaries.  It's a good point, though, and I'll leave it in comments
  in the file.}
%  \naive ~wins on \svdone with the exception to the 5 outliers.
The dashed line in Figures~\ref{fig:wpe_vs_mase_best}
and~\ref{fig:wpe_vs_mase_all} is a useful empirical guideline for
knowing when a model is not well-matched to the task at hand: a $MASE$
score that is more than {\color{red}[[what's the equation for that
      line, Ryan?]]} times the WPE of the corresponding signal is an
indication that that time series has more predictive structure than
that forecast model can capture and exploit.
